Galactic Chemical Evolution Models Favor an Extended Type Ia Supernova Delay-Time Distribution

### 2024-07-09 src/tex/ms.tex
[showyourwork] overleaf sync

\documentclass[twocolumn,twocolappendix,linenumbers]{aastex631}
% \documentclass[twocolumn,twocolappendix]{aastex631}

@@

Galactic chemical evolution (GCE) studies seek to explain the observed distribution of metals throughout the Milky Way Galaxy. \citet{Tinsley1979-StellarLifetimes} made a compelling case that the non-solar \aFe\footnote{
    In standard bracket notation, $[X/Y]\equiv \log_{10}(X/Y) - \log_{10}(X/Y)_{\odot}$. In this paper we will use \aFe and [O/Fe] interchangeably, although observational studies will often use a combination of $\alpha$-elements to calculate \aFe.
} ratios seen by, e.g., \citet{Wallerstein1962-GDwarfAbundances} were caused by different stellar lifetimes for the contributors of the Fe-peak elements than the $\alpha$-elements. Type Ia supernovae (SNe Ia), the thermonuclear explosions of carbon-oxygen white dwarfs (WDs), are responsible for a majority of the Fe produced in the Galaxy \citep{Matteucci1986-SupernovaEnrichment}; meanwhile, core collapse supernovae (CCSNe), the explosions of massive stars, produce the $\alpha$-elements (e.g., O and Mg) in addition to a smaller fraction of Fe. SNe Ia are delayed by $\sim{\bf 0.04}-10$ Gyr after star formation events, as evidenced by observations {\bf in both star-forming and} elliptical galaxies \citep[e.g.,][]{Maza1976-SNStatistics}. This delayed enrichment leads to a decrease in \aFe with increasing [Fe/H] \citep{Matteucci1986-SupernovaEnrichment}. Therefore, the relative abundances of the $\alpha$-elements and Fe as a function of stellar age trace the balance of SN rates over time.

@@

As a result of the uncertainty regarding SN Ia progenitors, theoretical models have yet to converge on a single prediction for the DTD. For the DD channel, assumptions about the distribution of {\bf WD} separations and the rate of gravitational wave inspiral suggest a broad $\sim t^{-1}$ DTD at long delay times ($\gtrsim 1$ Gyr), but at short delays ($\lesssim 1$ Gyr) the rate is limited by the need to produce two WDs {\bf \citep[see][]{Greggio2005-AnalyticalRates,Maoz2014-Review}}. Triple or higher-order progenitor systems could also produce a $t^{-1}$ DTD \citep{Fang2018-QuadrupleSystems,Rajamuthukumar2023-TripleEvolution}. The DTD which would result from the SD channel depends greatly on the assumptions of binary population synthesis, but in general is expected to cover a narrower range of delay times and may feature a steep exponential cutoff at the long end \citep[e.g.,][]{Greggio2005-AnalyticalRates}.

Surveys of SNe Ia can constrain the DTD by comparing the observed rate of SNe Ia to their host galaxy parameters \citep[e.g.,][]{Mannucci2005-SNRate,Heringer2019-FieldGalaxyDTD} or inferred star formation histories \citep[SFHs; e.g.,][]{Maoz2012-SloanIIDTD}, measuring SN Ia rates in galaxy clusters \citep[e.g.,][]{Maoz2010-ClusterDTD}, or comparing the volumetric SN Ia rate to the cosmic SFH as a whole \citep[e.g.,][]{Graur2014-VolumetricSNIaRates,Strolger2020-ExponentialDTD}. Early studies, which had limited sample sizes, produced unimodal \citep{Strolger2004-SNIaProgenitors} or bimodal \citep{Mannucci2006-TwoPopulations} DTDs where the majority of SNe Ia explode within a relatively narrow range of delay times. More recent studies have recovered broader DTD functions, with many converging on a declining power-law of $\sim t^{-1}$ {\bf \citep[e.g.,][]{Graur2013-IaRateVsMass,Graur2015-UnifiedExplanation,Maoz2017-CosmicDTD,Castrillo2021-DTD,Wiseman2021-DESRates}}, though there is some evidence for a steeper slope in galaxy clusters \citep{Maoz2017-CosmicDTD,FriedmannMaoz2018-ClusterDTD}. It is especially difficult to constrain the DTD for short delay times {\bf \citep{MaozMannucci2012-SNeIaReview,Rodney2014-PromptSNeIa}} because of the need for SN Ia rates at long lookback times and uncertainties in the age estimates of stellar populations.

@@

A few studies have investigated different DTDs in one-zone chemical evolution models, but comparisons to abundance data have been limited to the solar neighborhood \citep[e.g.,][]{Andrews2017-ChemicalEvolution,Palicio2023-AnalyticDTD}. \citet{Matteucci2009-DTDModels} compared five DTDs in a multi-zone GCE model and found that the agreement with observations worsens if the fraction of prompt {\bf ($\bf t\lesssim 100$ Myr)} SNe Ia is either too high or too low, but they were similarly limited by the available data for the solar neighborhood. \citet{Poulhazan2018-PrecisionPollution} found that the prompt component of the DTD affects the peak and width of the \aFe distribution in a cosmological smoothed-particle hydrodynamics simulation, but their simulation was not designed to reproduce the parameters of the Milky Way. The current era of large spectroscopic surveys such as the Apache Point Observatory Galactic Evolution Experiment \citep[APOGEE;][]{Majewski2017-APOGEE} and the ongoing Milky Way Mapper \citep{Kollmeier2017-SDSS-V} has made abundances across the Milky Way disk available for comparison to more sophisticated GCE models.

@@

For simplicity and easier comparison to the results of \citetalias{Johnson2021-Migration}, we focus our analysis on O and Fe, representing the $\alpha$ and Fe-peak elements, respectively. Both elements are produced by CCSNe. \vice adopts the instantaneous recycling approximation for CCSNe, so the equation which governs CCSN enrichment as a function of star formation for some element $x$ is simply
\begin{equation}
    \dot M_x^{\rm CC}{\bf (t)} = y_x^{\rm CC} \dot M_\star{\bf (t)}
    \label{eq:ccsn-enrichment}
\end{equation}
where $y_x^{\rm CC}$ is the CCSN yield of element $x$ per unit mass of star formation, and $\dot M_\star$ is the SFR. Following \citetalias{Johnson2021-Migration}, who in turn adopt their CCSN yields from \citet{ChieffiLimongi2004-CCSNYields} and \citet{LimongiChieffi2006-CCSNYields}, we adopt $y_{\rm O}^{\rm CC}=0.015$ and $y_{\rm Fe}^{\rm CC}=0.0012$. The primary effect of these yields is to set the low-[Fe/H] ``plateau'' in [O/Fe] which represents pure CCSN enrichment. The chosen yields for this paper produce a plateau at ${\rm [O/Fe]}=0.45$; see \citet{Weinberg2023-CCSNYield} for more discussion on the effect of the CCSN yields on chemical evolution.

Following the formalism of \citet{Weinberg2017-ChemicalEquilibrium}, the rate of Fe contribution to the ISM from SNe Ia is 
\begin{equation}
    \dot M_{\rm Fe}^{\rm Ia}{\bf (t)} = y_{\rm Fe}^{\rm Ia} \langle \dot M_\star\rangle_{\rm Ia} {\bf (t)}\\
\end{equation}
where $\langle \dot M_\star\rangle_{\rm Ia}{\bf (t)}$ is the time-averaged SFR weighted by the DTD {\bf at time $t$} and $y_{\rm Fe}^{\rm Ia}$ is the Fe yield of SNe Ia. \citet{Weinberg2017-ChemicalEquilibrium} show in their Appendix A that
\begin{equation}
    \langle \dot M_\star\rangle_{\rm Ia} \equiv \frac{\int_0^t \dot M_\star(t')R_{\rm Ia}(t-t')dt'}{\int_{t_D}^{t_{\rm max}} R_{\rm Ia}(t')dt'}
    \label{eq:weighted-sfr}
\end{equation}
where $R_{\rm Ia}$ is the DTD in units of ${\rm M}_\odot^{-1}\,{\rm yr}^{-1}$, $t_D$ is the minimum SN Ia delay time, and $t_{\rm max}$ is the lifetime of the disk. The denominator of Equation \ref{eq:weighted-sfr} is therefore equal to $N_{\rm Ia}/M_\star$, the total number of SNe Ia per M$_{\odot}$ of stars formed.

The yield $y_{\rm Fe}^{\rm Ia}$ measures the mass of Fe produced by SNe Ia over the full duration of the DTD, which can be expressed as:
\begin{equation}
    y_{\rm Fe}^{\rm Ia} = m_{\rm Fe}^{\rm Ia} \int_{t_D}^{t_{\rm max}} R_{\rm Ia}(t')dt' = m_{\rm Fe}^{\rm Ia} \frac{N_{\rm Ia}}{M_\star},
    \label{eq:dtd-integral}
\end{equation}
where $m_{\rm Fe}^{\rm Ia}$ is the average mass of Fe produced by a single SN Ia, and $N_{\rm Ia}/M_\star=2.2\pm1\times10^{-3}\,{\rm M}_\odot^{-1}$ is the average number of SNe Ia per mass of stars formed \citep{MaozMannucci2012-SNeIaReview}. Adjusting the value of $y_{\rm Fe}^{\rm Ia}$ primarily affects the end point of chemical evolution tracks in [O/Fe]--[Fe/H] space. Following \citetalias{Johnson2021-Migration}, we adopt $y_{\rm Fe}^{\rm Ia}=0.00214$. This yield is originally adapted from the W70 model of \citet{Iwamoto1999-SNIaYields}, but it is increased slightly so that the inside-out SFH produces stars with ${\rm [O/Fe]}\approx 0.0$ by the end of the model. {\bf The overall scale of the yields is inconsequential: a lower value of $y_{\rm Fe}^{\rm Ia}$ would produce similar results if compensated with a lower outflow mass-loading factor $\eta\equiv \dot M_{\rm out}/\dot M_\star$ (this is the yield-outflow degeneracy; see \citealt{Weinberg2023-CCSNYield}, \citealt{Sandford2024-StrongOutflows}, and appendix B of \citealt{Johnson2023-DwarfGalaxyArchaeology}).} \citet{Palla2021-SNIaYield} studied the effect of different SN Ia yields on GCE models in detail.

@@

We explore five different functional forms for the DTD: a two-population model, a single power-law, an exponential, a broken power-law with an initially flat plateau, and a model computed from triple-system dynamics. We also investigate one or two useful variations of the input parameters for each functional form. Figure \ref{fig:dtds} presents a selection of these DTDs, and Table \ref{tab:dtds} summarizes {\bf the parameters and median delay times ($t_{\rm med}$) for all of our DTDs}. We use simple forms rather than simulated physical or analytic models of SNe Ia for the sake of decreased computational time and easier interpretation of the model predictions. Physically-motivated models of the DTD must contend with many unknown or poorly-constrained parameters, so our simplified forms have the advantage of reducing the number of free parameters. In Appendix \ref{app:analytic-dtds}, we show that a few of our simple forms adequately approximate the more complete analytic models of \citet{Greggio2005-AnalyticalRates}.

{\bf In this subsection, we present functional forms of each DTD in terms of a function $f_{\rm Ia}$ that has units of ${\rm Gyr}^{-1}$ and defines the shape of the DTD for $t\ge t_D$ as}
\begin{equation}
    R_{\rm Ia}(t) = 
    \begin{cases}
        \frac{N_{\rm Ia}}{M_\star}
        \frac{f_{\rm Ia}(t)}{\int_{t_D}^{t_{\rm max}} f_{\rm Ia}(t') dt'}, & t \ge t_D \\
        0 & t < t_D.
    \end{cases}
    \label{eq:dtd-function}
\end{equation}
{\bf Note that the denominator in Equation \ref{eq:dtd-function} normalizes the DTD.}

@@

\paragraph{Power-law} A single power law with slope $\alpha$:
\begin{equation}
    f_{\rm Ia}^{\rm plaw}(t) = (t/1\,\rm{Gyr})^\alpha
    \label{eq:powerlaw-dtd}
\end{equation}
A declining power-law with $\alpha\sim-1$ {\bf \citep{Totani2008-DTD}} arises from typical assumptions about the distribution of post-common envelope separations and the rate of gravitational wave inspiral \citep[see Section 3.5 from][]{Maoz2014-Review}. It is therefore a commonly assumed DTD in GCE studies (e.g., \citealt{Rybizki2017-Chempy}; \citetalias{Johnson2021-Migration}; \citealt{Weinberg2023-CCSNYield}). Additionally, the observational evidence for a power-law DTD is strong. \citet{Maoz2017-CosmicDTD} obtained a DTD with $\alpha=-1.07\pm0.09$ based on volumetric rates and an assumed cosmic SFH for field galaxies in redshift range $0\leq z\leq 2.25$. \citet{Wiseman2021-DESRates} obtained a similar slope of $\alpha=-1.13\pm0.05$ for field galaxies in the redshift range $0.2<z<0.6$. {\bf \citet{Heringer2019-FieldGalaxyDTD} used a SFH-independent method to constrain the DTD for field galaxies within $0.01<z<0.2$ and found a larger value of $\alpha=-1.34^{+0.19}_{-0.17}$.} For galaxy clusters, \citet{Maoz2017-CosmicDTD} found a steeper DTD slope of $\alpha=-1.39^{+0.32}_{-0.05}$, {\bf and \citet{FriedmannMaoz2018-ClusterDTD} found a similar slope of $\alpha=-1.3^{+0.23}_{-0.16}$; however, a re-analysis by \citet{FreundlichMaoz2021-ClusterDTD} revealed no significant difference between the cluster and field galaxy DTDs}. In this paper, we investigate the cases $\alpha=-1.1$ and $\alpha=-1.4$.

\paragraph{Exponential} An exponentially declining DTD with timescale $\tau$:
\begin{equation}
    f_{\rm Ia}^{\rm exp}(t) = e^{-t/\tau}.
    \label{eq:exponential-dtd}
\end{equation}
This model allows analytic solutions to the abundances as a function of time for some SFHs, making it a popular choice {\bf \citep[e.g.,][]{Weinberg2017-ChemicalEquilibrium,Pantoni2019-AnalyticSolutions,Palicio2023-AnalyticDTD}}. \citet{Schonrich2009-RadialMixing} and \citet{Weinberg2017-ChemicalEquilibrium} both assumed an exponential DTD with a timescale $\tau=1.5$ Gyr. However, observational support for an exponential DTD {\bf is scarce}. \citet{Strolger2020-ExponentialDTD}, fitting to the cosmic SFH and SFHs from field galaxies, found a range of exponential-like solutions with timescales $\sim 1.5 - 6$ Gyr.
In this paper, we investigate timescales $\tau=1.5$ and 3 Gyr. {\bf We show in Appendix \ref{app:analytic-dtds} that an exponential DTD with $\tau=1.5$ Gyr is an adequate approximation for the analytic SD DTD from \citet{Greggio2005-AnalyticalRates}.}

\paragraph{Plateau} A modification of the power-law in which the DTD ``plateaus'' for a duration $W$ before declining:
\begin{equation}
    f_{\rm Ia}^{\rm plat}(t) =
    \begin{cases}
        1, & t < W \\
        (t/W)^\alpha, & t \ge W.
    \end{cases}
    \label{eq:plateau-dtd}
\end{equation}
Our primary motivation is to consider a model which matches observations at delay times beyond a few Gyr, where the DTD is best constrained, but with a smaller fraction of prompt ($\lesssim 100$ Myr) SNe Ia than the single power law. {\bf This form has been used for the DTD of neutron star mergers \citep{Simonetti2019-NeutronStarDTD}, but to our knowledge it has not been considered for SNe Ia in previous GCE models.} We show in Appendix \ref{app:analytic-dtds} that this form can approximate the more complicated analytic DD DTDs from \citet{Greggio2005-AnalyticalRates}. We investigate the cases $W=0.3$ Gyr and $W=1$ Gyr, taking $\alpha=-1.1$ for all plateau models.

@@

{\bf There are many models for the DTD which have been proposed in the literature, and an exhaustive test of every one is infeasible. In particular, we do not consider a Gaussian DTD \citep[e.g.,][]{Strolger2004-SNIaProgenitors}, in which most SNe Ia explode with delay times close to $\sim 3$ Gyr and there are very few prompt events. Evidence for this form came from SN Ia rate measurements for $z>1$, but re-analysis of the data revealed large errors due to the small sample size, extinction corrections, and uncertainties in the SFHs \citep[e.g.,][]{Forster2006-SNIaConstraints,Greggio2008-CosmicMix}. Additionally, \citet{Mannucci2006-TwoPopulations} found that such a DTD fails to reproduce the observed dependence of the SN Ia rate on galaxy color, and there is evidence that at least some SNe Ia must be prompt in order to explain observed rates in spiral galaxies \citep[e.g.,][]{Mannucci2005-SNRate,ScannapiecoBildsten2005-SNIaRate}. GCE studies have found that a Gaussian DTD over-produces high-[O/Fe] and low-[Fe/H] stars \citep{Matteucci2009-DTDModels,Palicio2023-AnalyticDTD}.}

@@

The inside-out and late-burst models are run in \vice's ``star formation mode,'' where the SFR surface density $\dot\Sigma_\star$ is prescribed along with the star formation efficiency (SFE) timescale $\tau_\star\equiv \Sigma_g/\dot\Sigma_\star$. The remaining quantities, infall rate surface density $\dot\Sigma_{\rm in}$ and gas surface density $\Sigma_g$, are calculated from the specified quantities {\bf assuming the star formation law adopted by \citetalias{Johnson2021-Migration} (see their Equation 14). More specifically, at each timestep the infall rate is calculated to fulfill the quantity of gas required to produce the specified SFR by the star formation law}. The latter two models are run in ``infall mode,'' where we specify $\dot\Sigma_{\rm in}$ {\bf and} $\tau_\star$. {\bf The initial gas mass is zero for all models (including those run in star formation mode).} The mode in which \vice models are run makes no difference as a unique solution can always be obtained if two of the four parametric forms are specified.

The SFH is normalized such that the model predicts {\bf a total stellar mass of $(5.17\pm1.11)\times10^{10} M_\odot$ \citep{LicquiaNewman2015-StellarMass} and the stellar surface density gradient reported by \citet[][see Appendix B of \citetalias{Johnson2021-Migration}]{BlandHawthornGerhard2016-MilkyWayReview}}.
We present an overview of the four SFHs in Figure \ref{fig:sfhs}, and we discuss them individually here.

@@

\subsection{\bf DTD Parameters: Slope, Timescale, and Plateau Width}
\label{sec:onezone-dtd-params}

@@

\subsection{\bf Qualitative Comparisons}
\label{sec:discussion-scores}

>>>>

\documentclass[twocolumn,twocolappendix]{aastex631}

@@

Galactic chemical evolution (GCE) studies seek to explain the observed distribution of metals throughout the Milky Way Galaxy. \citet{Tinsley1979-StellarLifetimes} made a compelling case that the non-solar \aFe\footnote{
    In standard bracket notation, $[X/Y]\equiv \log_{10}(X/Y) - \log_{10}(X/Y)_{\odot}$. In this paper we will use \aFe and [O/Fe] interchangeably, although observational studies will often use a combination of $\alpha$-elements to calculate \aFe.
} ratios seen by, e.g., \citet{Wallerstein1962-GDwarfAbundances} were caused by different stellar lifetimes for the contributors of the Fe-peak elements than the $\alpha$-elements. Type Ia supernovae (SNe Ia), the thermonuclear explosions of carbon-oxygen white dwarfs (WDs), are responsible for a majority of the Fe produced in the Galaxy \citep{Matteucci1986-SupernovaEnrichment}; meanwhile, core collapse supernovae (CCSNe), the explosions of massive stars, produce the $\alpha$-elements (e.g., O and Mg) in addition to a smaller fraction of Fe. SNe Ia are delayed by $\sim 0.04-10$ Gyr after star formation events, as evidenced by observations in both star-forming and elliptical galaxies \citep[e.g.,][]{Maza1976-SNStatistics}. This delayed enrichment leads to a decrease in \aFe with increasing [Fe/H] \citep{Matteucci1986-SupernovaEnrichment}. Therefore, the relative abundances of the $\alpha$-elements and Fe as a function of stellar age trace the balance of SN rates over time.

@@

As a result of the uncertainty regarding SN Ia progenitors, theoretical models have yet to converge on a single prediction for the DTD. For the DD channel, assumptions about the distribution of WD separations and the rate of gravitational wave inspiral suggest a broad $\sim t^{-1}$ DTD at long delay times ($\gtrsim 1$ Gyr), but at short delays ($\lesssim 1$ Gyr) the rate is limited by the need to produce two WDs \citep[see][]{Greggio2005-AnalyticalRates,Maoz2014-Review}. Triple or higher-order progenitor systems could also produce a $t^{-1}$ DTD \citep{Fang2018-QuadrupleSystems,Rajamuthukumar2023-TripleEvolution}. The DTD which would result from the SD channel depends greatly on the assumptions of binary population synthesis, but in general is expected to cover a narrower range of delay times and may feature a steep exponential cutoff at the long end \citep[e.g.,][]{Greggio2005-AnalyticalRates}.

Surveys of SNe Ia can constrain the DTD by comparing the observed rate of SNe Ia to their host galaxy parameters \citep[e.g.,][]{Mannucci2005-SNRate,Heringer2019-FieldGalaxyDTD} or inferred star formation histories \citep[SFHs; e.g.,][]{Maoz2012-SloanIIDTD}, measuring SN Ia rates in galaxy clusters \citep[e.g.,][]{Maoz2010-ClusterDTD}, or comparing the volumetric SN Ia rate to the cosmic SFH as a whole \citep[e.g.,][]{Graur2014-VolumetricSNIaRates,Strolger2020-ExponentialDTD}. Early studies, which had limited sample sizes, produced unimodal \citep{Strolger2004-SNIaProgenitors} or bimodal \citep{Mannucci2006-TwoPopulations} DTDs where the majority of SNe Ia explode within a relatively narrow range of delay times. More recent studies have recovered broader DTD functions, with many converging on a declining power-law of $\sim t^{-1}$ \citep[e.g.,][]{Graur2013-IaRateVsMass,Graur2015-UnifiedExplanation,Maoz2017-CosmicDTD,Castrillo2021-DTD,Wiseman2021-DESRates}, though there is some evidence for a steeper slope in galaxy clusters \citep{Maoz2017-CosmicDTD,FriedmannMaoz2018-ClusterDTD}. It is especially difficult to constrain the DTD for short delay times \citep{MaozMannucci2012-SNeIaReview,Rodney2014-PromptSNeIa} because of the need for SN Ia rates at long lookback times and uncertainties in the age estimates of stellar populations.

@@

A few studies have investigated different DTDs in one-zone chemical evolution models, but comparisons to abundance data have been limited to the solar neighborhood \citep[e.g.,][]{Andrews2017-ChemicalEvolution,Palicio2023-AnalyticDTD}. \citet{Matteucci2009-DTDModels} compared five DTDs in a multi-zone GCE model and found that the agreement with observations worsens if the fraction of prompt ($t\lesssim 100$ Myr) SNe Ia is either too high or too low, but they were similarly limited by the available data for the solar neighborhood. \citet{Poulhazan2018-PrecisionPollution} found that the prompt component of the DTD affects the peak and width of the \aFe distribution in a cosmological smoothed-particle hydrodynamics simulation, but their simulation was not designed to reproduce the parameters of the Milky Way. The current era of large spectroscopic surveys such as the Apache Point Observatory Galactic Evolution Experiment \citep[APOGEE;][]{Majewski2017-APOGEE} and the ongoing Milky Way Mapper \citep{Kollmeier2017-SDSS-V} has made abundances across the Milky Way disk available for comparison to more sophisticated GCE models.

@@

For simplicity and easier comparison to the results of \citetalias{Johnson2021-Migration}, we focus our analysis on O and Fe, representing the $\alpha$ and Fe-peak elements, respectively. Both elements are produced by CCSNe. \vice adopts the instantaneous recycling approximation for CCSNe, so the equation which governs CCSN enrichment as a function of star formation for some element $x$ is simply
\begin{equation}
    \dot M_x^{\rm CC}(t) = y_x^{\rm CC} \dot M_\star(t)
    \label{eq:ccsn-enrichment}
\end{equation}
where $y_x^{\rm CC}$ is the CCSN yield of element $x$ per unit mass of star formation, and $\dot M_\star$ is the SFR. Following \citetalias{Johnson2021-Migration}, who in turn adopt their CCSN yields from \citet{ChieffiLimongi2004-CCSNYields} and \citet{LimongiChieffi2006-CCSNYields}, we adopt $y_{\rm O}^{\rm CC}=0.015$ and $y_{\rm Fe}^{\rm CC}=0.0012$. The primary effect of these yields is to set the low-[Fe/H] ``plateau'' in [O/Fe] which represents pure CCSN enrichment. The chosen yields for this paper produce a plateau at ${\rm [O/Fe]}=0.45$; see \citet{Weinberg2023-CCSNYield} for more discussion on the effect of the CCSN yields on chemical evolution.

Following the formalism of \citet{Weinberg2017-ChemicalEquilibrium}, the rate of Fe contribution to the ISM from SNe Ia is 
\begin{equation}
    \dot M_{\rm Fe}^{\rm Ia}(t) = y_{\rm Fe}^{\rm Ia} \langle \dot M_\star\rangle_{\rm Ia}(t)\\
\end{equation}
where $\langle \dot M_\star\rangle_{\rm Ia}(t)$ is the time-averaged SFR weighted by the DTD at time $t$ and $y_{\rm Fe}^{\rm Ia}$ is the Fe yield of SNe Ia. \citet{Weinberg2017-ChemicalEquilibrium} show in their Appendix A that
\begin{equation}
    \langle \dot M_\star\rangle_{\rm Ia} \equiv \frac{\int_0^t \dot M_\star(t')R_{\rm Ia}(t-t')dt'}{\int_{t_D}^{t_{\rm max}} R_{\rm Ia}(t')dt'}
    \label{eq:weighted-sfr}
\end{equation}
where $R_{\rm Ia}$ is the DTD in units of ${\rm M}_\odot^{-1}\,{\rm yr}^{-1}$, $t_D$ is the minimum SN Ia delay time, and $t_{\rm max}$ is the lifetime of the disk. The denominator of Equation \ref{eq:weighted-sfr} is therefore equal to $N_{\rm Ia}/M_\star$, the total number of SNe Ia per M$_{\odot}$ of stars formed.

The yield $y_{\rm Fe}^{\rm Ia}$ measures the mass of Fe produced by SNe Ia over the full duration of the DTD, which can be expressed as:
\begin{equation}
    y_{\rm Fe}^{\rm Ia} = m_{\rm Fe}^{\rm Ia} \int_{t_D}^{t_{\rm max}} R_{\rm Ia}(t')dt' = m_{\rm Fe}^{\rm Ia} \frac{N_{\rm Ia}}{M_\star},
    \label{eq:dtd-integral}
\end{equation}
where $m_{\rm Fe}^{\rm Ia}$ is the average mass of Fe produced by a single SN Ia, and $N_{\rm Ia}/M_\star=2.2\pm1\times10^{-3}\,{\rm M}_\odot^{-1}$ is the average number of SNe Ia per mass of stars formed \citep{MaozMannucci2012-SNeIaReview}. Adjusting the value of $y_{\rm Fe}^{\rm Ia}$ primarily affects the end point of chemical evolution tracks in [O/Fe]--[Fe/H] space. Following \citetalias{Johnson2021-Migration}, we adopt $y_{\rm Fe}^{\rm Ia}=0.00214$. This yield is originally adapted from the W70 model of \citet{Iwamoto1999-SNIaYields}, but it is increased slightly so that the inside-out SFH produces stars with ${\rm [O/Fe]}\approx 0.0$ by the end of the model. The overall scale of the yields is inconsequential: a lower value of $y_{\rm Fe}^{\rm Ia}$ would produce similar results if compensated with a lower outflow mass-loading factor $\eta\equiv \dot M_{\rm out}/\dot M_\star$ (this is the yield-outflow degeneracy; see \citealt{Weinberg2023-CCSNYield}, \citealt{Sandford2024-StrongOutflows}, and appendix B of \citealt{Johnson2023-DwarfGalaxyArchaeology}). \citet{Palla2021-SNIaYield} studied the effect of different SN Ia yields on GCE models in detail.

@@

We explore five different functional forms for the DTD: a two-population model, a single power-law, an exponential, a broken power-law with an initially flat plateau, and a model computed from triple-system dynamics. We also investigate one or two useful variations of the input parameters for each functional form. Figure \ref{fig:dtds} presents a selection of these DTDs, and Table \ref{tab:dtds} summarizes the parameters and median delay times ($t_{\rm med}$) for all of our DTDs. We use simple forms rather than simulated physical or analytic models of SNe Ia for the sake of decreased computational time and easier interpretation of the model predictions. Physically-motivated models of the DTD must contend with many unknown or poorly-constrained parameters, so our simplified forms have the advantage of reducing the number of free parameters. In Appendix \ref{app:analytic-dtds}, we show that a few of our simple forms adequately approximate the more complete analytic models of \citet{Greggio2005-AnalyticalRates}.

In this subsection, we present functional forms of each DTD in terms of a function $f_{\rm Ia}$ that has units of ${\rm Gyr}^{-1}$ and defines the shape of the DTD for $t\ge t_D$ as
\begin{equation}
    R_{\rm Ia}(t) = 
    \begin{cases}
        \frac{N_{\rm Ia}}{M_\star}
        \frac{f_{\rm Ia}(t)}{\int_{t_D}^{t_{\rm max}} f_{\rm Ia}(t') dt'}, & t \ge t_D \\
        0 & t < t_D.
    \end{cases}
    \label{eq:dtd-function}
\end{equation}
Note that the denominator in Equation \ref{eq:dtd-function} normalizes the DTD.

@@

\paragraph{Power-law} A single power law with slope $\alpha$:
\begin{equation}
    f_{\rm Ia}^{\rm plaw}(t) = (t/1\,\rm{Gyr})^\alpha
    \label{eq:powerlaw-dtd}
\end{equation}
A declining power-law with $\alpha\sim-1$ \citep{Totani2008-DTD} arises from typical assumptions about the distribution of post-common envelope separations and the rate of gravitational wave inspiral \citep[see Section 3.5 from][]{Maoz2014-Review}. It is therefore a commonly assumed DTD in GCE studies (e.g., \citealt{Rybizki2017-Chempy}; \citetalias{Johnson2021-Migration}; \citealt{Weinberg2023-CCSNYield}). Additionally, the observational evidence for a power-law DTD is strong. \citet{Maoz2017-CosmicDTD} obtained a DTD with $\alpha=-1.07\pm0.09$ based on volumetric rates and an assumed cosmic SFH for field galaxies in redshift range $0\leq z\leq 2.25$. \citet{Wiseman2021-DESRates} obtained a similar slope of $\alpha=-1.13\pm0.05$ for field galaxies in the redshift range $0.2<z<0.6$. \citet{Heringer2019-FieldGalaxyDTD} used a SFH-independent method to constrain the DTD for field galaxies within $0.01<z<0.2$ and found a larger value of $\alpha=-1.34^{+0.19}_{-0.17}$. For galaxy clusters, \citet{Maoz2017-CosmicDTD} found a steeper DTD slope of $\alpha=-1.39^{+0.32}_{-0.05}$, and \citet{FriedmannMaoz2018-ClusterDTD} found a similar slope of $\alpha=-1.3^{+0.23}_{-0.16}$; however, a re-analysis by \citet{FreundlichMaoz2021-ClusterDTD} revealed no significant difference between the cluster and field galaxy DTDs. In this paper, we investigate the cases $\alpha=-1.1$ and $\alpha=-1.4$.

\paragraph{Exponential} An exponentially declining DTD with timescale $\tau$:
\begin{equation}
    f_{\rm Ia}^{\rm exp}(t) = e^{-t/\tau}.
    \label{eq:exponential-dtd}
\end{equation}
This model allows analytic solutions to the abundances as a function of time for some SFHs, making it a popular choice \citep[e.g.,][]{Weinberg2017-ChemicalEquilibrium,Pantoni2019-AnalyticSolutions,Palicio2023-AnalyticDTD}. \citet{Schonrich2009-RadialMixing} and \citet{Weinberg2017-ChemicalEquilibrium} both assumed an exponential DTD with a timescale $\tau=1.5$ Gyr. However, observational support for an exponential DTD is scarce. \citet{Strolger2020-ExponentialDTD}, fitting to the cosmic SFH and SFHs from field galaxies, found a range of exponential-like solutions with timescales $\sim 1.5 - 6$ Gyr.
In this paper, we investigate timescales $\tau=1.5$ and 3 Gyr. We show in Appendix \ref{app:analytic-dtds} that an exponential DTD with $\tau=1.5$ Gyr is an adequate approximation for the analytic SD DTD from \citet{Greggio2005-AnalyticalRates}.

\paragraph{Plateau} A modification of the power-law in which the DTD ``plateaus'' for a duration $W$ before declining:
\begin{equation}
    f_{\rm Ia}^{\rm plat}(t) =
    \begin{cases}
        1, & t < W \\
        (t/W)^\alpha, & t \ge W.
    \end{cases}
    \label{eq:plateau-dtd}
\end{equation}
Our primary motivation is to consider a model which matches observations at delay times beyond a few Gyr, where the DTD is best constrained, but with a smaller fraction of prompt ($\lesssim 100$ Myr) SNe Ia than the single power law. This form has been used for the DTD of neutron star mergers \citep{Simonetti2019-NeutronStarDTD}, but to our knowledge it has not been considered for SNe Ia in previous GCE models. We show in Appendix \ref{app:analytic-dtds} that this form can approximate the more complicated analytic DD DTDs from \citet{Greggio2005-AnalyticalRates}. We investigate the cases $W=0.3$ Gyr and $W=1$ Gyr, taking $\alpha=-1.1$ for all plateau models.

@@

There are many models for the DTD which have been proposed in the literature, and an exhaustive test of every one is infeasible. In particular, we do not consider a Gaussian DTD \citep[e.g.,][]{Strolger2004-SNIaProgenitors}, in which most SNe Ia explode with delay times close to $\sim 3$ Gyr and there are very few prompt events. Evidence for this form came from SN Ia rate measurements for $z>1$, but re-analysis of the data revealed large errors due to the small sample size, extinction corrections, and uncertainties in the SFHs \citep[e.g.,][]{Forster2006-SNIaConstraints,Greggio2008-CosmicMix}. Additionally, \citet{Mannucci2006-TwoPopulations} found that such a DTD fails to reproduce the observed dependence of the SN Ia rate on galaxy color, and there is evidence that at least some SNe Ia must be prompt in order to explain observed rates in spiral galaxies \citep[e.g.,][]{Mannucci2005-SNRate,ScannapiecoBildsten2005-SNIaRate}. GCE studies have found that a Gaussian DTD over-produces high-[O/Fe] and low-[Fe/H] stars \citep{Matteucci2009-DTDModels,Palicio2023-AnalyticDTD}.

@@

The inside-out and late-burst models are run in \vice's ``star formation mode,'' where the SFR surface density $\dot\Sigma_\star$ is prescribed along with the star formation efficiency (SFE) timescale $\tau_\star\equiv \Sigma_g/\dot\Sigma_\star$. The remaining quantities, infall rate surface density $\dot\Sigma_{\rm in}$ and gas surface density $\Sigma_g$, are calculated from the specified quantities assuming the star formation law adopted by \citetalias{Johnson2021-Migration} (see their Equation 14). More specifically, at each timestep the infall rate is calculated to fulfill the quantity of gas required to produce the specified SFR by the star formation law. The latter two models are run in ``infall mode,'' where we specify $\dot\Sigma_{\rm in}$ and $\tau_\star$. The initial gas mass is zero for all models (including those run in star formation mode). The mode in which \vice models are run makes no difference as a unique solution can always be obtained if two of the four parametric forms are specified.

The SFH is normalized such that the model predicts a total stellar mass of $(5.17\pm1.11)\times10^{10} M_\odot$ \citep{LicquiaNewman2015-StellarMass} and the stellar surface density gradient reported by \citet[][see Appendix B of \citetalias{Johnson2021-Migration}]{BlandHawthornGerhard2016-MilkyWayReview}.
We present an overview of the four SFHs in Figure \ref{fig:sfhs}, and we discuss them individually here.

@@

\subsection{DTD Parameters: Slope, Timescale, and Plateau Width}
\label{sec:onezone-dtd-params}

@@

\subsection{Qualitative Comparisons}
\label{sec:discussion-scores}
